---
title: "visualization"
subtitle: "Part I"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
#library(hexbin)
## hexbin is needed for running geom_hex()
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Scatterplots

create the first scatterplot!

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point()
```

new approach, same plot.

```{r}
weather_df %>% 
  ## we can do filter, select here
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```


save and edit a plot object
```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot
## save a plot into some object and print it later
```


## Advanced scatterplot

start with the same one and make it fancy!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE)
## remove the shadow of se
```

waht about the `aes` placement?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name)) +
  geom_smooth()
## we only get one blue line across the plot. the reason is that the color statement is within the first geom_point statement, but it will not influence the second one.
```

Let's facet some things!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
## "." means don't create rows, "~ name" means use name to create columns

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)


weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .2, size = .2) +
  geom_smooth(se = FALSE, size = 2) +
  facet_grid(. ~ name)
## alpha = .5 means 50% transparency


weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) + 
  geom_point() +
  geom_smooth(se = FALSE, size = 2) +
  facet_grid(. ~ name)
## alpha could equal to a variable
```

Let's combine some elements and try a new plot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## small notes

How many geoms have to exist?

You can have whatever geoms you want.

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

You can use a neat geom!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))
  #geom_hex()
  #geom_bin2d()
  #geom_density2d()
```


## univariate plots

Histogram!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

can we add color?
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, color = name)) +
  geom_histogram(position = "dodge")
# dodge is used to make the bars not overlapped with each other

weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")

weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

a new geometry!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .4, adjust = .5)
# density is used to smooth the data, but we lose some information, but easy to compare
# adjust means to show more details, like binwidth in histogram
```

what about box plot?

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

Trendy plots
```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = .5) +
#more intuitive than boxplots
  stat_summary(fun = "median")
```

Ridge plot _ the most popular one in 2017

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
# more easier to see when we have many categories
# need to load ggridges package to run geom_ridges
```