---
title: "visualization"
subtitle: "Part II"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## Load the weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Remember this plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5)
```

## labels

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'temperature plot',
    x = "Minimun daily temperature (C)",
    x = "maximum daily temperature (C)",
    caption = "data from rnoaa package; temperature in 2017"
  )
```


## scales

start with the same plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ยบ C", "0", "15")) +
  scale_y_continuous(
    #trans = "sqrt"
    ## to transfore the y axis
    position = "right"
  )
```

Look at color scales

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'temperature plot',
    x = "Minimun daily temperature (C)",
    x = "maximum daily temperature (C)",
    caption = "data from rnoaa package; temperature in 2017"
  ) +
  scale_color_hue(
    name = "Location",
    ## to change the name of the index table
    h = c(200, 400))



weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'temperature plot',
    x = "Minimun daily temperature (C)",
    x = "maximum daily temperature (C)",
    caption = "data from rnoaa package; temperature in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) # the default is continuous, use this statement to let R know this data is discrete
```

## Themes

Shift the legend

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'temperature plot',
    x = "Minimun daily temperature (C)",
    x = "maximum daily temperature (C)",
    caption = "data from rnoaa package; temperature in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
  theme(legend.position = "bottom") #legend means the index table
```

change the overall theme
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  labs(
    title = 'temperature plot',
    x = "Minimun daily temperature (C)",
    x = "maximum daily temperature (C)",
    caption = "data from rnoaa package; temperature in 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE) +
 # theme_bw()
 # theme_minimal()
 # theme_classoc()
ggthemes::theme_economist() # Lots of themes in ggthemes package

## theme statement will wipe up any statement above and do the plot as the theme's rule. SO ORDER MATTERS!
```

## setting options

```{r eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal) + theme(legend.position = "bottom")

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_color_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

## run this in the beginning will determine the theme of every plot in the document
```


## data orgs in `geom`

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
# if don't specify the dataset, the default dataset will be waikiki, so we can specify the data in the geom() statement for the specific geom
```

## `patchwork`

remember faceting?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .5) + 
  facet_grid(. ~ name)
```

What happens when you want multipanel plots but can't facet?

```{r}
## patchwork is used to put multiple plots into one panel
tmax_tmin_p = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
# Go to the patchwork webpage to see more info
```

## Data manipulation

Control your factors

```{r}
weather_df %>%
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```


What if I wanted densities for tmin and tmax simultaneously?

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
# this is a data manupulation issue rather than a ggplot issue
```

## revisit the pups

data from the FAS study
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  # starts_with("pd_") can help to select all pd_'s
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) %>% 
  ## to reorder the "pd_"s
  # mutate(outcome = forcats::fct_relevel(outcome, "pd_ears", "pd_pivot", "pd_walk", "pd_eyes")) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```



```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```







